
\section{Education, Outreach, Result from Prior Support}
 
\noindent
\textbf{Education and outreach.}
Implementing and maintaining a large software system such as PC requires significant engineering effort. 
Rice undergraduates have supplied a significant amount of the
engineering effort so far, and we will continue this going forward.
A significant portion of 
engineering effort will be provided by Rice undergraduates, through directed, independent studies, as well as, if possible, 
paid positions (we plan to apply for an REU supplement if funded).  
We will place a special effort on recruiting women.
In addition, funded by the salary support requested, PI Jermaine will continue his efforts to mentor local high school students in
computer science research.  
Currently, Jermaine mentors one local female student from William P. Clements High School on a deep learning capstone
project.  
The goal 
is to mentor two high school students at all times during the lifetime of the project.
Those student will attend Jermaine's weekly research group meetings, and meet with PI Jermaine at least bi-weekly.

\vspace{5 pt}
\noindent
\textbf{Results from prior NSF support.}
The project closest to this one is \textit{
SimSQL: A Database System Supporting Implementation and Execution of Distributed Machine Learning Codes}; 
August 2014 -- September 2018.
Statistical machine learning is a commonly-applied framework for analyzing very large data sets. The SimSQL project (award 1409543) is concerned with performing the fundamental research necessary to make ML-in-the-database a mature technology. \textbf{Intellectual Merit:} We have developed new database programming languages for statistical ML \cite{jermaineBUDS}, considered how to add matrix and vector support to a database \cite{jermaineVector}, how to represent objects \cite{jermaineComplex}, how the database/user-defined function interface should change to support ML \cite{jermaineUDF}, how the complex objects necessary for ML should be represented in a database system \cite{jermaineExperimental}, and how optimization should be performed in such a system \cite{rusu2015workload}. \textbf{Broader Impact:} The project has made it easier to write ML codes and get them to run in a distributed environment, enabling a strong separation between the code and the database data processing algorithms, storage, hardware, and indexing. So far, we have released a prototype of the system, under the Apache license.

